Investigating Predictors for American Attitudes on Justifiability of Abortion

### 2022-06-21 Final/FinalReport.qmd
added conclusion + discussion section

\[paste here after finalizing the google doc\]

@@

@@

>>>>

The model that we fitted to predict abortion attitudes generally confirmed our initial research question that liberal ideology and attitudes towards other issues predicts abortion attitudes. Inference on the model (through both a confidence interval and hypothesis test) showed that having liberal political ideology, low belief in the importance of God, and belief in greater child autonomy generally predict a belief that abortion is more justified. However, holding liberal attitudes on certain other issues, namely believing that respect for authority is bad and not having pride in being American, do not necessarily predict greater justifiability of abortion.

Despite the model confirming our initial research question, we have reason to question the quality of the model, and whether the model accurately fits the data on abortion attitudes. The first reason for questioning our model’s quality is the low R-squared value, which indicates that the variance of our predictors may not greatly account for the variance in abortion attitudes. The second reason for questioning our model’s quality is that our model does not fit the linearity and constant variance conditions. Upon closer examination of the data, we can see that the distribution of our response variable was trimodal, with peaks at the extremes and in the middle. This likely led to the data having a large number of very different observations that had different predictor values, but the same `aj` score, and may be one explanation our model’s weaknesses.

Given the nature of public opinion on political issues, we are likely limited by the precision and accuracy of each observation in our dataset. Previous political science research on public opinion indicates that Americans’ political beliefs are inconsistent and volatile. For example, most Americans’ political beliefs have been found to be shaped by partisan and group loyalties, rather than ideology or any specific set of issue positions. Indeed, even for committed partisans, one individual’s issue positions are likely to to be scattered across the ideological spectrum, and may even change over short periods of time (Achen and Bartels, 2017). This inconsistency between partisanship, ideology, and issue positions may account for the non-normal distribution of abortion attitudes in the dataset and the model’s poor fit to the data. In addition, although the data are supposed to capture a nationally representative sample of the US population, we do not know if they are weighted by other factors that may represent important distinctions in the US population. This possibility of non-weighting may have influenced our model’s quality as well.

Future work may build on this analysis and correct some of its weaknesses. First, the data should be weighted by more relevant factors to ensure that the sample is as representative as possible of the US population. Relevant factors may include, but are not limited to, partisan affiliation, race, religion, socioeconomic status, and geographic area. In addition, it may be helpful to consider other predictors that may influence abortion attitudes, such as partisan affiliation and race. However, it is important to keep in mind that abortion is an incredibly divisive and complex issue, so it is not an easy task to precisely pinpoint any individual’s abortion attitude, no matter which predictors are used.

@@

Achen, C.H., & Bartels, L.M. (2017). Democracy for realists: Why elections do not produce responsive government. Princeton University Press.

@@

Miller, Steven V. (2016, Jul 31).  “Attitudes on the Justifiability of Abortion in the United States.”  GitHub.  https://github.com/svmiller/wvs-usa-abortion-attitudes
